\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2023-02-21 Latex/paper.tex
# last version
I'm aworking with the convergence order.

% custom macros and definitions
\newtheorem{theorem}{Theorem}[section] 
%%\newtheorem{acknowledgement}[theorem]{Acknowledgement} 
%%\newtheorem{algorithm}[theorem]{Algorithm} 
%%\newtheorem{axiom}[theorem]{Axiom} 
%%\newtheorem{case}[theorem]{Case} 
%%\newtheorem{claim}[theorem]{Claim} 
%%\newtheorem{conclusion}[theorem]{Conclusion} 
%%\newtheorem{condition}[theorem]{Condition} 
%%\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary} 
%%\newtheorem{criterion}[theorem]{Criterion} 
%%\newtheorem{defi}[theorem]{Definition} 
\newtheorem{example}[theorem]{Example} 
%%\newtheorem{exercise}[theorem]{Exercise} 
%%\newtheorem{hypothesis}[theorem]{Hypothesis} 
\newtheorem{lemma}[theorem]{Lemma} 
%%\newtheorem{notation}[theorem]{Notation} 
%%\newtheorem{problem}[theorem]{Problem} 
\newtheorem{proposition}[theorem]{Proposition} 
%%\newtheorem{question}[theorem]{Question} 
\newtheorem{remark}[theorem]{Remark} 
%%\newtheorem{solution}[theorem]{Solution} 
%%\newtheorem{summary}[theorem]{Summary} 
%%\newtheorem{Rem}[theorem]{Remark}

@@

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function and $h$ a non-negative process. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01}), that in this case (\ref{ois_forward_rate_curve})
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta_t dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Then with a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}. 

@@

\section*{Appendix}
\appendix
\renewcommand{\thesection}{\Alph{section}.\arabic{section}}
\section{Approximation of $\mathbb{E}_s^{t_p}\left(D_s x_{t_a}\right)$}
\label{expected_forward_measure_malliavin_derivative_x_t}
From (\ref{short_rate_cheyette}) is easy to show that
\begin{equation*}
x_{t_a} = \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) y_u du + \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(u,x_u,y_u) dW_u^{\mathbb{Q}}  
\end{equation*}
In order to get an approximation for $D_sX_{t_a}$ we must to avoid the recurrence in the Malliavin derivative of $X_{t_a}$. Follow, the ideas of Piterbarg Volume II, we can approximate $y_t$ the next way
\begin{equation*}
y_t \approx \int_{0}^{t} \exp\left(-2\int_{u}^{t} k_{u^{\prime}} du^{\prime} \right) \eta^{2}(u,x_0,y_0) du
\end{equation*}
Therefore, if we define $\bar{y}(t)= \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0)$, we have that
\begin{equation}\label{approximation_x_t_a}
X_{t_a} \approx \bar{X}(t_a):= \bar{x}_0(t_a)  + \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{y}(u) du + \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(u,\bar{x}(u),\bar{y}(u)) dW_u^{\mathbb{Q}}  
\end{equation}
With initial condition  $\bar{x}_{0}(t_a)$ is such that $S_{a,b}(t_a,\bar{x}_0(t_a), \bar{y}_{t_a}) = S_{a,b}(0)$ (see Piterbarg Volume II). Now, we must remember that 
\begin{equation*}
dW^{\mathbb{Q}^{t_p}} = dW^{\mathbb{Q}} + \bar{\nu}(t,t_p) dt 
\end{equation*}
where $\bar{\nu}(t,t_p) = \int_{t}^{t_p} \eta(s,\bar{x}_s,\bar{y}_s) ds $. Then, we have the next approximation of $\bar{x}_{t_a}$ under measure $\mathbb{Q}^{t_p}$
\begin{align*}
X_{t_a} &\approx \bar{x}_0  + \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{y}_s ds - \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{\nu}(s, t_p) \eta(s,\bar{x}_s,\bar{y}_s) ds +  \\
&+ \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right)\eta(s,\bar{x}_s,\bar{y}_s) dW_s^{\mathbb{Q}^{t_p}}  
\end{align*} 
Then, if we take $D_s$ on $\bar{X}_{t_a}$ we get that (esto hay que formalizarlo...)
\begin{equation}\label{approximation_D_s_x_t}
D_sX_{t_a} \approx  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(s,\bar{x}_0(t_a),\bar{y}(t_a))\bar{M}(s,t_a)
\end{equation}
where
\begin{align*}
\bar{M}(s,t_a) &= \exp\left(-\int_{s}^{t_a} \left( \frac{\left(\partial_x \beta(u,t_a,\bar{x}_0,\bar{y}_{t_a})\right)^{2}}{2} - \exp\left(-\int_{u}^{t_a}k_u du\right) \partial_x (\eta(u, \bar{x}_u, \bar{y}_{u}) \bar{\nu}(u,t_p))\right) du \right) \\ 
&\cdot\exp\left(\int_{s}^{t_0} \partial_x \beta(u,t_a,\bar{x}_0,\bar{y}_{t_a}) dW^{\mathbb{Q}}_u \right)
\end{align*}
With $\beta(u,t_a,x,y) = \exp\left(-\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) \partial_x \eta(u,x,y)$. Therefore, from previous equalities, we have that

@@

\section{Order of the approximation $\bar{y}_t$}
Under the assumption of $\eta(t,\cdot,\cdot)$ is Lipschitz and with first derive continuous, we have that

>>>>

% custom macros and definitions
\newtheorem{theorem}{Theorem}[section] 
%%\newtheorem{acknowledgement}[theorem]{Acknowledgement} 
%%\newtheorem{algorithm}[theorem]{Algorithm} 
%%\newtheorem{axiom}[theorem]{Axiom} 
%%\newtheorem{case}[theorem]{Case} 
%%\newtheorem{claim}[theorem]{Claim} 
%%\newtheorem{conclusion}[theorem]{Conclusion} 
%%\newtheorem{condition}[theorem]{Condition} 
%%\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary} 
%%\newtheorem{criterion}[theorem]{Criterion} 
%%\newtheorem{defi}[theorem]{Definition} 
\newtheorem{example}[theorem]{Example} 
%%\newtheorem{exercise}[theorem]{Exercise} 
\newtheorem{hypothesis}[theorem]{Hypothesis} 
\newtheorem{lemma}[theorem]{Lemma} 
%%\newtheorem{notation}[theorem]{Notation} 
%%\newtheorem{problem}[theorem]{Problem} 
\newtheorem{proposition}[theorem]{Proposition} 
%%\newtheorem{question}[theorem]{Question} 
\newtheorem{remark}[theorem]{Remark} 
%%\newtheorem{solution}[theorem]{Solution} 
%%\newtheorem{summary}[theorem]{Summary} 
%%\newtheorem{Rem}[theorem]{Remark}

@@

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function and $h$ a non-negative process. In addition
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Throught this paper, we assume the following hypotheses on $\eta(t,x,y)$.
\begin{hypothesis}\label{boundedness_volatility} 
The process $\eta_t$ is global Lipschitz and differentiable a.s. In addition, we will suppose that
\begin{align*}
a1 \leq \eta(t,x,y) \leq a2 \quad \forall (t,x,t) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+}. \\
|\eta(t,x_2,y_2) - \eta(t,x_2,y_2)| \leq C_{x,y}\left(|x_2 - x_1| + |y_2 - y_1| \right) \quad \forall (t,x,t) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+} 
\end{align*}
\end{hypothesis}
\begin{hypothesis}\label{boundedness_reversion} 
The next integrals are finite
\begin{align*}
I_{1,k}(0,t) &= \int_{0}^{t} \exp\left(- 2 \int_{s}^{t} k(s^{\prime})ds^{\prime}\right)  ds < \infty \\
I_{2,k}(0,t) &= \int_{0}^{t} \exp\left(- \int_{s}^{t} k(s^{\prime})ds^{\prime}\right) G(s,t) ds < \infty
\end{align*}
\end{hypothesis}
\begin{remark}
Notice that the above hypotheses have been chosen for the sake of simplicity, but they can be replaced by adequate integrability conditions.
\end{remark}
We must note, that under the hypothesis (\ref{boundedness_volatility}) $\partial_x \eta(t,x,y)$ and $\partial_y \eta(t,x,y)$ are bounded. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01} that
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta_t dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Now if we do a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0} 

@@

\section*{Appendix}
\appendix
\renewcommand{\thesection}{\Alph{section}.\arabic{section}}

\section{Order of the approximation $\bar{y}_t$}
Under the assumption of $\eta(t,\cdot,\cdot)$ is global Lipschitz and with first derive continuous, we have that
\begin{align*}
|\eta^{2}(t,x_t,y_t) - \eta^{2}(t,x_0,y_0)|&=|\eta(t,x_t,y_t) - \eta(t,x_0,y_0)||\eta(t,x_t,y_t) + \eta(t,x_0,y_0)| \nonumber \\ 
&\leq C_{x,y}\left(|x_t - x_0| + |y_t - y_0|\right) + 
\end{align*}

\section{Approximation of $\mathbb{E}_s^{t_p}\left(D_s x_{t_a}\right)$}
\label{expected_forward_measure_malliavin_derivative_x_t}
From (\ref{short_rate_cheyette}) is easy to show that
\begin{equation*}
x_{t_a} = \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) y_u du + \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(u,x_u,y_u) dW_u^{\mathbb{Q}}  
\end{equation*}
In order to get an approximation for $D_sX_{t_a}$ we must to avoid the recurrence in the Malliavin derivative of $X_{t_a}$. Follow, the ideas of Piterbarg Volume II, we can approximate $y_t$ the next way
\begin{equation*}
y_t \approx \int_{0}^{t} \exp\left(-2\int_{u}^{t} k_{u^{\prime}} du^{\prime} \right) \eta^{2}(u,x_0,y_0) du
\end{equation*}
Therefore, if we define $\bar{y}(t)= \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0)$, we have that
\begin{equation}\label{approximation_x_t_a}
X_{t_a} \approx \bar{X}(t_a):= \bar{x}_0(t_a)  + \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{y}(u) du + \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(u,\bar{x}(u),\bar{y}(u)) dW_u^{\mathbb{Q}}  
\end{equation}
With initial condition  $\bar{x}_{0}(t_a)$ is such that $S_{a,b}(t_a,\bar{x}_0(t_a), \bar{y}_{t_a}) = S_{a,b}(0)$ (see Piterbarg Volume II). Now, we must remember that 
\begin{equation*}
dW^{\mathbb{Q}^{t_p}} = dW^{\mathbb{Q}} + \bar{\nu}(t,t_p) dt 
\end{equation*}
where $\bar{\nu}(t,t_p) = \int_{t}^{t_p} \eta(s,\bar{x}_s,\bar{y}_s) ds $. Then, we have the next approximation of $\bar{x}_{t_a}$ under measure $\mathbb{Q}^{t_p}$
\begin{align*}
X_{t_a} &\approx \bar{x}_0  + \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{y}_s ds - \int_{0}^{t_a} \exp\left(-\int_{s}^{t_a}k_u du\right) \bar{\nu}(s, t_p) \eta(s,\bar{x}_s,\bar{y}_s) ds +  \\
&+ \int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u du \right)\eta(s,\bar{x}_s,\bar{y}_s) dW_s^{\mathbb{Q}^{t_p}}  
\end{align*} 
Then, if we take $D_s$ on $\bar{X}_{t_a}$ we get that (esto hay que formalizarlo...)
\begin{equation}\label{approximation_D_s_x_t}
D_sX_{t_a} \approx  \exp\left(-\int_{s}^{t_a}k_u du \right) \eta(s,\bar{x}_0(t_a),\bar{y}(t_a))\bar{M}(s,t_a)
\end{equation}
where
\begin{align*}
\bar{M}(s,t_a) &= \exp\left(-\int_{s}^{t_a} \left( \frac{\left(\partial_x \beta(u,t_a,\bar{x}_0,\bar{y}_{t_a})\right)^{2}}{2} - \exp\left(-\int_{u}^{t_a}k_u du\right) \partial_x (\eta(u, \bar{x}_u, \bar{y}_{u}) \bar{\nu}(u,t_p))\right) du \right) \\ 
&\cdot\exp\left(\int_{s}^{t_0} \partial_x \beta(u,t_a,\bar{x}_0,\bar{y}_{t_a}) dW^{\mathbb{Q}}_u \right)
\end{align*}
With $\beta(u,t_a,x,y) = \exp\left(-\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) \partial_x \eta(u,x,y)$. Therefore, from previous equalities, we have that

@@
